---
title: "MLM_US_Survey"
author: "CH"
date: "06/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here) # find relative path
library(readr) # read dataset
library(tidyverse) # data wrangling
library(dplyr) # data wrangling
library(nlme) # multi-level regression
library(lme4) # multi-level regression
library(lmerTest) # multi-level regression
library(dfoptim)
library(optimx)

# dataset ####
data <- here::here("data/US_data.csv")

US_data <- readr::read_csv(data) %>%
  as_tibble() 

US_data$time <- factor(US_data$time)
US_data$unexp <- factor(US_data$unexp)
US_data$vote <- factor(US_data$vote)
US_data <- US_data %>%
  mutate(vote = case_when(
  vote=='1' ~ 'Democrats',
  vote=='2' ~ 'Republicans',
  TRUE ~ 'Other'))

US_data <- US_data %>%
  mutate(unexp = case_when(
    unexp == "0" ~ "Expected",
    unexp == "1" ~ "Unexpected",
    unexp == "2" ~ "Neither expected nor unexpected"))

US_data <- US_data %>%
  mutate(gend = case_when(
    gend == "1" ~ "Female",
    gend == "2" ~ "Male",
    gend == "3" ~ "Non-binary"
  ))

US_data <- US_data %>%
  mutate(ethn = case_when(
    ethn == "1" ~ "Asian-American",
    ethn == "2" ~ "Black or African American",
    ethn == "3" ~ "Hispanic or Latino American",
    ethn == "4" ~ "Native American",
    ethn == "5" ~ "White or European American",
    ethn == "6" ~ "Other"
  ))

US_data <- US_data %>%
  mutate(degree =case_when(
    degree == "1" ~ "No degree",
    degree == "2" ~ "High School",
    degree == "3" ~ "Bachelor´s degree",
    degree == "4" ~ "Master´s degree",
    degree == "5" ~ "Doctoral degree",
    degree == "6" ~ "Professional degree",
    degree == "7" ~ "Other"
  ))

US_data <- US_data %>%
  mutate(employ = case_when(
    employ == "1" ~ "Working - paid employee",
    employ == "2" ~ "Working - self-employed",
    employ == "3" ~ "Not working - temporary layoff from a job",
    employ == "4" ~ "Not working - looking for a job",
    employ == "5" ~ "Not working - retired",
    employ == "6" ~ "Not working - disabled",
    employ == "7" ~ "Not working - student",
    employ == "8" ~ "Other"
  ))

US_data <- US_data %>%
  mutate(income = case_when(
    income == "1" ~ "$0 - $10,000",
    income == "2" ~ "$10,000 - $20,000",
    income == "3" ~ "$20,000 - $30,000",
    income == "4" ~ "$30,000 - $40,000",
    income == "5" ~ "$40,000 - $50,000",
    income == "6" ~ "over $50,000"
  ))

US <- US_data %>%
  mutate(cn = (cn_1 + cn_2 + cn_3 + cn_4 + cn_5)/5,
        secident = (secident_1 + secident_2 + secident_3 + secident_4 + secident_5 +
        secident_6 + secident_7 + secident_8 + secident_9 + secident_10)/10,
        ingroup_satis = (secident_4 + secident_5 + secident_6 + secident_7)/4,
        xeno = (xeno_1 + xeno_2 + xeno_3 + xeno_4)/4, 
        joy = (joy_1 + joy_2)/2,
        eff = (eff_1 + eff_2)/2) %>%
  as.tibble()
US

US$time <- factor(US$time)
US$unexp <- factor(US$unexp)
US$vote <- factor(US$vote)

# credits go to Jenny Terry for providing advice and code (https://github.com/jenny-terry)

# select variables for MLM analyses and filter Republicans
US_mlm <- US %>%
  dplyr::select(., id, time, vote, unexp, cn, secident, sd, joy, eff, fc_1_o_agr, fc_2_o_agr, cor_values, cor_health, sign, donate, share) %>%
  dplyr::filter(., vote == "Republicans")
head(US_mlm)

# convert character variable to factors

US_mlm$id <- factor(US_mlm$id)

# standardize variables which are on different scales (than 1 to 7)
US_mlm$sd = scale(US_mlm$sd)
US_mlm$fc_1_o_agr = scale(US_mlm$fc_1_o_agr)
US_mlm$fc_2_o_agr = scale(US_mlm$fc_2_o_agr)
US_mlm$cor_health = scale(US_mlm$cor_health)
US_mlm$cor_values = scale(US_mlm$cor_values)
US_mlm$donate = scale(US_mlm$donate)

# mlm is chosen to account for (repeated measures within-participant) dependency: the scores measured are nested within the individual with makes the individual (id) = level 2, and the values level 1 --> we need to center by id ("group" center)

# group mean center level 1 predictors (note: fc_1_o_agr, fc_2_o_agr, joy and eff are also DVs and will be added in their non-centered form when used as a DV)

US_mlm <- US_mlm %>%
  dplyr::group_by(id) %>%
  dplyr::mutate(gmc_cn = mean(cn), gmc_secident = mean(secident), gmc_joy = mean(joy), gmc_eff = mean(eff))

allFit(show.meth.tab = TRUE)
```

# cn --> fc_1_o_agr
```{r}
# unconditional model (grand mean fc_1_o_agr across all individuals and time points)
cn_fc_1_o_agr.mod <- lmer(fc_1_o_agr ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(cn_fc_1_o_agr.mod)
lme4::confint.merMod(cn_fc_1_o_agr.mod)

# unconditional growth model (variance across individuals and time)
cn_fc_1_o_agr.mod2 <- lmer(fc_1_o_agr ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_fc_1_o_agr.mod2)
lme4::confint.merMod(cn_fc_1_o_agr.mod2)
cn_fc_1_o_agr.mod2_all <- allFit(cn_fc_1_o_agr.mod2)

# conditional growth model (inclusion of covariates time & secident and interaction between time and cn as time-variant covariate that could possibly change with time)
cn_fc_1_o_agr.mod3 <- lmer(fc_1_o_agr ~ time + gmc_secident + time*gmc_cn + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_fc_1_o_agr.mod3)
lme4::confint.merMod(cn_fc_1_o_agr.mod3)

anova(cn_fc_1_o_agr.mod, cn_fc_1_o_agr.mod2, cn_fc_1_o_agr.mod3)

###




# conditional growth model (inclusion of secident)
cn_fc_1_o_agr.mod4 <- lmer(fc_1_o_agr ~ time + gmc_secident + time*gmc_cn + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_fc_1_o_agr.mod4)
lme4::confint.merMod(cn_fc_1_o_agr.mod4)
```

# cn --> fc_2_o_agr
```{r}
# unconditional model (grand mean fc_1_o_agr across all individuals and time points)
cn_fc_2_o_agr.mod <- lmer(fc_2_o_agr ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(cn_fc_2_o_agr.mod)
lme4::confint.merMod(cn_fc_2_o_agr.mod)

# unconditional growth model (variance across individuals and time)
cn_fc_2_o_agr.mod2 <- lmer(fc_2_o_agr ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_fc_2_o_agr.mod2)
lme4::confint.merMod(cn_fc_2_o_agr.mod2)

# conditional growth model (inclusion of TVC (= time-variant covariates) predictors that could possibly change as a function of time)
cn_fc_2_o_agr.mod3 <- lmer(fc_2_o_agr ~ time + gmc_secident + time*gmc_cn + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_fc_2_o_agr.mod3)
lme4::confint.merMod(cn_fc_2_o_agr.mod3)

anova(cn_fc_2_o_agr.mod, cn_fc_2_o_agr.mod2, cn_fc_2_o_agr.mod3)

###




# conditional growth model (inclusion of secident)
cn_fc_2_o_agr.mod4 <- lmer(fc_2_o_agr ~ gmc_cn + gmc_secident + time*gmc_cn + time*gmc_secident + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_fc_2_o_agr.mod4)
lme4::confint.merMod(cn_fc_2_o_agr.mod4)
```

# cn --> signing
```{r}
# unconditional model (logisitc regression)
cn_sign.mod <- glmer(sign ~ 1 + (1|id), US_mlm, family = "binomial")
summary(cn_sign.mod)



# unconditional growth model
cn_sign.mod2 <- glmer(sign ~ time + (time|id), US_mlm, family = "binomial", control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_sign.mod2)



# conditional growth model
cn_sign.mod3 <- glmer(sign ~ time + gmc_secident + time*gmc_cn + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_sign.mod3)
lme4::confint.merMod(cn_sign.mod3)
or.glmer::(US_mlm, cn_sign.mod3, complex)


anova(cn_sign.mod, cn_sign.mod2, cn_sign.mod3)

### 




# conditional growth model secident added
cn_sign.mod4 <- glmer(sign ~ gmc_cn + gmc_secident + gmc_cn*time + gmc_secident*time + (time|id), US_mlm, family = "binomial", control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_sign.mod4)
cc_cn_sign.mod4 <- confint(cn_sign.mod4, parm = "beta_")
ctab_cn_sign.mod4 <- cbind(est=fixef(cn_sign.mod4),cc_cn_sign.mod4)
rtab_cn_sign.mod4 <- exp(ctab_cn_sign.mod4)
print(rtab_cn_sign.mod4, digits = 3)
```

# fc_1_o_agr --> joy
```{r}
# unconditional model
fc_1_o_agr_joy.mod <- lmer(joy ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(fc_1_o_agr_joy.mod)
lme4::confint.merMod(fc_1_o_agr_joy.mod)

# unconditional growth model
fc_1_o_agr_joy.mod2 <- lmer(joy ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_1_o_agr_joy.mod2)
lme4::confint.merMod(fc_1_o_agr_joy.mod2)
cc_fc_1_o_agr_joy.mod2 <- confint(fc_1_o_agr_joy.mod2, parm = "beta_")
ctab_fc_1_o_agr_joy.mod2 <- cbind(est=fixef(fc_1_o_agr_joy.mod2),cc_fc_1_o_agr_joy.mod2)
rtab_fc_1_o_agr_joy.mod2 <- exp(ctab_fc_1_o_agr_joy.mod2)
print(rtab_fc_1_o_agr_joy.mod2, digits = 3)

# conditional growth model
fc_1_o_agr_joy.mod3 <- lmer(joy ~ time + time*fc_1_o_agr + (time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_1_o_agr_joy.mod3)
lme4::confint.merMod(fc_1_o_agr_joy.mod3)
cc_fc_1_o_agr_joy.mod3 <- confint(fc_1_o_agr_joy.mod3, parm = "beta_")
ctab_fc_1_o_agr_joy.mod3 <- cbind(est=fixef(fc_1_o_agr_joy.mod3),cc_fc_1_o_agr_joy.mod3)
rtab_fc_1_o_agr_joy.mod3 <- exp(ctab_fc_1_o_agr_joy.mod3)
print(rtab_fc_1_o_agr_joy.mod3, digits = 3)

anova(fc_1_o_agr_joy.mod, fc_1_o_agr_joy.mod2, fc_1_o_agr_joy.mod3)
```

# fc_2_o_agr --> joy
```{r}
# unconditional model
fc_2_o_agr_joy.mod <- lmer(joy ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(fc_2_o_agr_joy.mod)
lme4::confint.merMod(fc_2_o_agr_joy.mod)

# unconditional growth model
fc_2_o_agr_joy.mod2 <- lmer(joy ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_2_o_agr_joy.mod2)
lme4::confint.merMod(fc_2_o_agr_joy.mod2)

# conditional growth model
fc_2_o_agr_joy.mod3 <- lmer(joy ~ time + time*fc_2_o_agr + (time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_2_o_agr_joy.mod3)
lme4::confint.merMod(fc_2_o_agr_joy.mod3)
cc_fc_2_o_agr_joy.mod3 <- confint(fc_2_o_agr_joy.mod3, parm = "beta_")
ctab_fc_2_o_agr_joy.mod3 <- cbind(est=fixef(fc_2_o_agr_joy.mod3),cc_fc_2_o_agr_joy.mod3)
rtab_fc_2_o_agr_joy.mod3 <- exp(ctab_fc_2_o_agr_joy.mod3)
print(rtab_fc_2_o_agr_joy.mod3, digits = 3)

anova(fc_2_o_agr_joy.mod, fc_2_o_agr_joy.mod2, fc_2_o_agr_joy.mod3)
```

# fc_1_o_agr --> eff
```{r}
# unconditional model
fc_1_o_agr_eff.mod <- lmer(eff ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(fc_1_o_agr_eff.mod)
lme4::confint.merMod(fc_1_o_agr_eff.mod)

# unconditional growth model
fc_1_o_agr_eff.mod2 <- lmer(eff ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_1_o_agr_eff.mod2)
lme4::confint.merMod(fc_1_o_agr_eff.mod2)
cc_fc_1_o_agr_eff.mod2 <- confint(fc_1_o_agr_eff.mod2, parm = "beta_")
ctab_fc_1_o_agr_eff.mod2 <- cbind(est=fixef(fc_1_o_agr_eff.mod2),cc_fc_1_o_agr_eff.mod2)
rtab_fc_1_o_agr_eff.mod2 <- exp(ctab_fc_1_o_agr_eff.mod2)
print(rtab_fc_1_o_agr_eff.mod2, digits = 3)

# conditional growth model
fc_1_o_agr_eff.mod3 <- lmer(eff ~ time + time*fc_1_o_agr + (time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_1_o_agr_eff.mod3)
lme4::confint.merMod(fc_1_o_agr_eff.mod3)
cc_fc_1_o_agr_eff.mod3 <- confint(fc_1_o_agr_eff.mod3, parm = "beta_")
ctab_fc_1_o_agr_eff.mod3 <- cbind(est=fixef(fc_1_o_agr_eff.mod3),cc_fc_1_o_agr_eff.mod3)
rtab_fc_1_o_agr_eff.mod3 <- exp(ctab_fc_1_o_agr_eff.mod3)
print(rtab_fc_1_o_agr_eff.mod3, digits = 3)

anova(fc_1_o_agr_eff.mod, fc_1_o_agr_eff.mod2, fc_1_o_agr_eff.mod3)
```

# fc_2_o_agr --> eff
```{r}
fc_2_o_agr_eff.mod <- lmer(eff ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(fc_2_o_agr_eff.mod)
lme4::confint.merMod(fc_2_o_agr_eff.mod)

# unconditional growth model
fc_2_o_agr_eff.mod2 <- lmer(eff ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_2_o_agr_eff.mod2)
lme4::confint.merMod(fc_2_o_agr_eff.mod2)

# conditional growth model
fc_2_o_agr_eff.mod3 <- lmer(eff ~ time + time*fc_2_o_agr + (time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_2_o_agr_eff.mod3)
lme4::confint.merMod(fc_2_o_agr_eff.mod3)
cc_fc_2_o_agr_eff.mod3 <- confint(fc_2_o_agr_eff.mod3, parm = "beta_")
ctab_fc_2_o_agr_eff.mod3 <- cbind(est=fixef(fc_2_o_agr_eff.mod3),cc_fc_2_o_agr_eff.mod3)
rtab_fc_2_o_agr_eff.mod3 <- exp(ctab_fc_2_o_agr_eff.mod3)
print(rtab_fc_2_o_agr_eff.mod3, digits = 3)

anova(fc_2_o_agr_eff.mod, fc_2_o_agr_eff.mod2, fc_2_o_agr_eff.mod3)
```

# fc_1_o_agr --> signing
```{r}
# unconditional model (logisitc regression)
fc_1_o_agr_sign.mod <- glmer(sign ~ 1 + (1|id), US_mlm, family = "binomial")
summary(fc_1_o_agr_sign.mod)
# CI and OR
cc_fc_1_o_agr_sign.mod <- confint(fc_1_o_agr_sign.mod, parm = "beta_")
ctab_fc_1_o_agr_sign.mod <- cbind(est=fixef(fc_1_o_agr_sign.mod),cc_fc_1_o_agr_sign.mod)
rtab_fc_1_o_agr_sign.mod <- exp(ctab_fc_1_o_agr_sign.mod)
print(rtab_fc_1_o_agr_sign.mod, digits = 3)

# unconditional growth model
fc_1_o_agr_sign.mod2 <- glmer(sign ~ time + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_1_o_agr_sign.mod2)
cc_fc_1_o_agr_sign.mod2 <- confint(fc_1_o_agr_sign.mod2, parm = "beta_")
ctab_fc_1_o_agr_sign.mod2 <- cbind(est=fixef(fc_1_o_agr_sign.mod2),cc_fc_1_o_agr_sign.mod2)
rtab_fc_1_o_agr_sign.mod2 <- exp(ctab_fc_1_o_agr_sign.mod2)
print(rtab_fc_1_o_agr_sign.mod2, digits = 3)
lme4::confint.merMod(fc_1_o_agr_sign.mod2)

# conditional growth model
fc_1_o_agr_sign.mod3 <- glmer(sign ~ time + time*fc_1_o_agr + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_1_o_agr_sign.mod3)
cc_fc_1_o_agr_sign.mod3 <- confint(fc_1_o_agr_sign.mod3, parm = "beta_")
ctab_fc_1_o_agr_sign.mod3 <- cbind(est=fixef(fc_1_o_agr_sign.mod3),cc_fc_1_o_agr_sign.mod3)
rtab_fc_1_o_agr_sign.mod3 <- exp(ctab_fc_1_o_agr_sign.mod3)
print(rtab_fc_1_o_agr_sign.mod3, digits = 3)
lme4::confint.merMod(fc_1_o_agr_sign.mod3)

anova(fc_1_o_agr_sign.mod, fc_1_o_agr_sign.mod2, fc_1_o_agr_sign.mod3)
```
# fc_2_o_agr --> signing
```{r}
# unconditional model (logisitc regression)
fc_2_o_agr_sign.mod <- glmer(sign ~ 1 + (1|id), US_mlm, family = "binomial")
summary(fc_2_o_agr_sign.mod)
# CI and OR
cc_fc_2_o_agr_sign.mod <- confint(fc_1_o_agr_sign.mod, parm = "beta_")
ctab_fc_2_o_agr_sign.mod <- cbind(est=fixef(fc_2_o_agr_sign.mod),cc_fc_2_o_agr_sign.mod)
rtab_fc_2_o_agr_sign.mod <- exp(ctab_fc_2_o_agr_sign.mod)
print(rtab_fc_2_o_agr_sign.mod, digits = 3)

# unconditional growth model
fc_2_o_agr_sign.mod2 <- glmer(sign ~ time + (time|id), US_mlm, family = "binomial", control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_2_o_agr_sign.mod2)
cc_fc_2_o_agr_sign.mod2 <- confint(fc_2_o_agr_sign.mod2, parm = "beta_")
ctab_fc_2_o_agr_sign.mod2 <- cbind(est=fixef(fc_2_o_agr_sign.mod2),cc_fc_2_o_agr_sign.mod2)
rtab_fc_2_o_agr_sign.mod2 <- exp(ctab_fc_2_o_agr_sign.mod2)
print(rtab_fc_2_o_agr_sign.mod2, digits = 3)

# conditional growth model
fc_2_o_agr_sign.mod3 <- glmer(sign ~ time + fc_2_o_agr*time + (time|id), US_mlm, family = "binomial", control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(fc_2_o_agr_sign.mod3)
cc_fc_2_o_agr_sign.mod3 <- confint(fc_2_o_agr_sign.mod3, parm = "beta_")
ctab_fc_2_o_agr_sign.mod3 <- cbind(est=fixef(fc_2_o_agr_sign.mod3),cc_fc_2_o_agr_sign.mod3)
rtab_fc_2_o_agr_sign.mod3 <- exp(ctab_fc_2_o_agr_sign.mod3)
print(rtab_fc_2_o_agr_sign.mod3, digits = 3)

anova(fc_2_o_agr_sign.mod, fc_2_o_agr_sign.mod2, fc_2_o_agr_sign.mod3)
```

# joy --> signing
```{r}
# unconditional model (logisitc regression)
joy_sign.mod <- glmer(sign ~ 1 + (1|id), US_mlm, family = "binomial")
summary(joy_sign.mod)
# CI and OR
cc_joy_sign.mod <- confint(joy_sign.mod, parm = "beta_")
ctab_joy_sign.mod <- cbind(est=fixef(joy_sign.mod),cc_joy_sign.mod)
rtab_joy_sign.mod <- exp(ctab_joy_sign.mod)
print(rtab_joy_sign.mod, digits = 3)

# unconditional growth model
joy_sign.mod2 <- glmer(sign ~ time + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(joy_sign.mod2)
cc_joy_sign.mod2 <- confint(joy_sign.mod2, parm = "beta_")
ctab_joy_sign.mod2 <- cbind(est=fixef(joy_sign.mod2),cc_joy_sign.mod2)
rtab_joy_sign.mod2 <- exp(ctab_joy_sign.mod2)
print(rtab_joy_sign.mod2, digits = 3)

# conditional growth model
joy_sign.mod3 <- glmer(sign ~ time + gmc_joy*time + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(joy_sign.mod3)
cc_joy_sign.mod3 <- confint(joy_sign.mod3, parm = "beta_")
ctab_joy_sign.mod3 <- cbind(est=fixef(joy_sign.mod3),cc_joy_sign.mod3)
rtab_joy_sign.mod3 <- exp(ctab_joy_sign.mod3)
print(rtab_joy_sign.mod3, digits = 3)

anova(joy_sign.mod, joy_sign.mod2, joy_sign.mod3)
```

# eff --> signing
```{r}
# unconditional model (logisitc regression)
eff_sign.mod <- glmer(sign ~ 1 + (1|id), US_mlm, family = "binomial")
summary(eff_sign.mod)
# CI and OR
cc_eff_sign.mod <- confint(eff_sign.mod, parm = "beta_")
ctab_eff_sign.mod <- cbind(est=fixef(eff_sign.mod),cc_eff_sign.mod)
rtab_eff_sign.mod <- exp(ctab_eff_sign.mod)
print(rtab_eff_sign.mod, digits = 3)

# unconditional growth model
eff_sign.mod2 <- glmer(sign ~ time + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(eff_sign.mod2)
cc_eff_sign.mod2 <- confint(eff_sign.mod2, parm = "beta_")
ctab_eff_sign.mod2 <- cbind(est=fixef(eff_sign.mod2),cc_eff_sign.mod2)
rtab_eff_sign.mod2 <- exp(ctab_eff_sign.mod2)
print(rtab_eff_sign.mod2, digits = 3)

# conditional growth model
eff_sign.mod3 <- glmer(sign ~ time + gmc_eff*time + (time|id), US_mlm, family = "binomial", control = glmerControl(check.nobs.vs.nRE = "ignore"))
summary(eff_sign.mod3)
cc_eff_sign.mod3 <- confint(eff_sign.mod3, parm = "beta_")
ctab_eff_sign.mod3 <- cbind(est=fixef(eff_sign.mod3),cc_eff_sign.mod3)
rtab_eff_sign.mod3 <- exp(ctab_eff_sign.mod3)
print(rtab_eff_sign.mod3, digits = 3)

anova(eff_sign.mod, eff_sign.mod2, eff_sign.mod3)
```

# cn --> joy
```{r}
# unconditional model (grand mean fc_1_o_agr across all individuals and time points)
cn_joy.mod <- lmer(joy ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(cn_joy.mod)
lme4::confint.merMod(cn_joy.mod)

# unconditional growth model (variance across individuals and time)
cn_joy.mod2 <- lmer(joy ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_joy.mod2)
lme4::confint.merMod(cn_joy.mod2)

# conditional growth model (inclusion of secident)
cn_joy.mod3 <- lmer(joy ~ time + gmc_secident + time*gmc_cn + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_joy.mod3)
lme4::confint.merMod(cn_joy.mod3)

anova(cn_joy.mod, cn_joy.mod2, cn_joy.mod3)

###




# conditional growth model (inclusion of TVC (= time-variant covariates) predictors that could possibly change as a function of time)
cn_joy.mod4 <- lmer(joy ~ gmc_cn + time*gmc_cn + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_joy.mod3)
lme4::confint.merMod(cn_joy.mod4)

cn_joy.mod5 <- lmer(joy ~ time + secident + time*gmc_cn + (time|id), US_mlm, REML = FALSE,control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_joy.mod5)
```

# cn --> eff
```{r}
# unconditional model (grand mean fc_1_o_agr across all individuals and time points)
cn_eff.mod <- lmer(eff ~ 1 + (1|id), data = US_mlm, REML = FALSE)
summary(cn_eff.mod)
lme4::confint.merMod(cn_eff.mod)

# unconditional growth model (variance across individuals and time)
cn_eff.mod2 <- lmer(eff ~ time + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_eff.mod2)
lme4::confint.merMod(cn_eff.mod2)

# conditional growth model (inclusion of TVC (= time-variant covariates) predictors that could possibly change as a function of time)
cn_eff.mod3 <- lmer(eff ~ time + gmc_secident + time*gmc_cn + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_eff.mod3)
lme4::confint.merMod(cn_eff.mod3)

anova(cn_eff.mod, cn_eff.mod2, cn_eff.mod3)

###



# conditional growth model (inclusion of secident)
cn_eff.mod4 <- lmer(eff ~ gmc_cn + gmc_secident + time*gmc_cn + time*gmc_secident + (time|id), data = US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(cn_eff.mod4)
lme4::confint.merMod(cn_eff.mod4)
```













# OLD

# cn --> fc_1_o_agr step 1: unconditional model
```{r}
# intercept only
mod1 <- lmerTest::lmer(fc_1_o_agr ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(mod1)
lme4::confint.merMod(mod1)

# adding predictor cn
mod2 <- lmerTest::lmer(fc_1_o_agr ~ cn + (1|id), US_mlm, REML = FALSE)
summary(mod2)
lme4::confint.merMod(mod2)

# adding "covariates" secident and time
mod3 <- lmerTest::lmer(fc_1_o_agr ~ cn + secident + time + (1|id), US_mlm, REML = FALSE)
summary(mod3)
lme4::confint.merMod(mod3)

# model fit comparison
anova(mod1, mod2, mod3)
```

# cn --> fc_1_o_agr step 2: adding group means (level 1 & 2 effects), group (id) specific differences around the overall mean
```{r}
# adding group mean cn
mod4 <- lmerTest::lmer(fc_1_o_agr ~ cn + gmc_cn + time + (1|id), US_mlm, REML = FALSE)
summary(mod4)
lme4::confint.merMod(mod4)

# adding group mean secident and time
mod5 <- lmerTest::lmer(fc_1_o_agr ~ cn + gmc_cn + secident + gmc_secident + time + (1|id), US_mlm, REML = FALSE)
summary(mod5)
lme4::confint.merMod(mod5)

anova(mod1, mod2, mod3, mod4, mod5)
```

# cn --> fc_1_o_agr step 3: random slopes for time
```{r}
mod6 <- lmerTest::lmer(fc_1_o_agr ~ cn + gmc_cn + secident + gmc_secident + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(mod6)
lme4::confint.merMod(mod6)

# failed to converge
allFit(show.meth.tab = TRUE)
mod6_all <- allFit(mod6)
```

# cn --> fc_2_o_agr step 1
```{r}
mod1a <- lmerTest::lmer(fc_2_o_agr ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(mod1a)
lme4::confint.merMod(mod1a)

# adding predictor cn
mod2a <- lmerTest::lmer(fc_2_o_agr ~ cn + (1|id), US_mlm, REML = FALSE)
summary(mod2a)
lme4::confint.merMod(mod2a)

# adding "covariates" secident and time
mod3a <- lmerTest::lmer(fc_2_o_agr ~ cn + secident + time + (1|id), US_mlm, REML = FALSE)
summary(mod3a)
lme4::confint.merMod(mod3a)

# model fit comparison
anova(mod1a, mod2a, mod3a)
```

# cn --> fc_2_o_agr step 2
```{r}
# adding group mean cn
mod4a <- lmerTest::lmer(fc_2_o_agr ~ cn + gmc_cn + (1|id), US_mlm, REML = FALSE)
summary(mod4a)
lme4::confint.merMod(mod4a)

# adding group mean secident and time
mod5a <- lmerTest::lmer(fc_2_o_agr ~ cn + gmc_cn + secident + gmc_secident + time + (1|id), US_mlm, REML = FALSE)
summary(mod5a)
lme4::confint.merMod(mod5a)

anova(mod1a, mod2a, mod3a, mod4a, mod5a)
```

# cn --> fc_2_o_agr step 3
```{r}
mod6a <- lmerTest::lmer(fc_2_o_agr ~ cn + gmc_cn + secident + gmc_secident + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(mod6a)
lme4::confint.merMod(mod6a)

# failed to converge
allFit(show.meth.tab = TRUE)
mod6a_all <- allFit(mod6a)
```

# cn --> behavioural intention
```{r}
sign.mod <- glmer(sign ~ 1 + (1|id), US_mlm, family = "binomial")
summary(sign.mod)

sign.mod2 <- glmer(sign ~ cn + (1|id), US_mlm, family = "binomial")
summary(sign.mod2)

sign.mod3 <- glmer(sign ~ cn + secident + time + (1|id), US_mlm, family = "binomial")
summary(sign.mod3)

sign.mod4 <- glmer(sign ~ cn + gmc_cn + secident + gmc_secident + time + (1|id), US_mlm, family = "binomial")
summary(sign.mod4)

sign.mod5 <- glmer(sign ~ cn + gmc_cn + secident + gmc_secident + time + (1|id) + (0+time|id), US_mlm, family = "binomial")
summary(sign.mod5)

anova(sign.mod, sign.mod2, sign.mod3, sign.mod4, sign.mod5)
```

# fc_1_o_agr --> joy step 1
```{r}
# intercept only
mod1b <- lmerTest::lmer(joy ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(mod1b)
lme4::confint.merMod(mod1b)

# adding predictor fc_1_o_agr and time
mod2b <- lmerTest::lmer(joy ~ fc_1_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod2b)
lme4::confint.merMod(mod2b)

# model fit comparison
anova(mod1b, mod2b)
```

# fc_1_o_agr --> joy step 2
```{r}
# adding group mean centered fc_1_o_agr
mod3b <- lmerTest::lmer(joy ~ fc_1_o_agr + gmc_fc_1_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod3b)
lme4::confint.merMod(mod3b)

anova(mod1b, mod2b, mod3b)
```

# fc_1_o_agr --> joy step 3
```{r}
# random slopes time
mod4b <- lmerTest::lmer(joy ~ fc_1_o_agr + gmc_fc_1_o_agr + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(mod4b)
lme4::confint.merMod(mod4b)

# failed to converge
allFit(show.meth.tab = TRUE)
mod4b_all <- allFit(mod4b)
```

# fc_2_o_agr --> joy step 1
```{r}
# intercept only
mod1c <- lmerTest::lmer(joy ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(mod1b)
lme4::confint.merMod(mod1c)

# adding predictor fc_2_o_agr and time
mod2c <- lmerTest::lmer(joy ~ fc_2_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod2c)
lme4::confint.merMod(mod2c)

# model fit comparison
anova(mod1c, mod2c)
```

# fc_2_o_agr --> joy step 2
```{r}
# adding group mean centered fc_2_o_agr
mod3c <- lmerTest::lmer(joy ~ fc_2_o_agr + gmc_fc_2_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod3c)
lme4::confint.merMod(mod3c)

anova(mod1c, mod2c, mod3c)
```

# fc_2_o_agr --> joy step 3
```{r}
# random slopes time
mod4c <- lmerTest::lmer(joy ~ fc_2_o_agr + gmc_fc_2_o_agr + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(mod4c)
lme4::confint.merMod(mod4c)

# failed to converge
allFit(show.meth.tab = TRUE)
mod4c_all <- allFit(mod4c)

anova(mod1c, mod2c, mod3c, mod4c)
```


# fc_1_o_agr --> eff step 1
```{r}
# intercept only
mod1d <- lmerTest::lmer(eff ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(mod1d)
lme4::confint.merMod(mod1d)

# adding predictor fc_1_o_agr and time
mod2d <- lmerTest::lmer(eff ~ fc_1_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod2d)
lme4::confint.merMod(mod2d)

# model fit comparison
anova(mod1d, mod2d)
```

# fc_1_o_agr --> eff step 2
```{r}
# adding group mean centered fc_1_o_agr
mod3d <- lmerTest::lmer(eff ~ fc_1_o_agr + gmc_fc_1_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod3d)
lme4::confint.merMod(mod3d)

anova(mod1d, mod2d, mod3d)
```

# fc_1_o_agr --> eff step 3
```{r}
# random slopes time
mod4d <- lmerTest::lmer(eff ~ fc_1_o_agr + gmc_fc_1_o_agr + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(mod4d)
lme4::confint.merMod(mod4d)

# failed to converge
allFit(show.meth.tab = TRUE)
mod4d_all <- allFit(mod4d)

anova(mod1d, mod2d, mod3d, mod4d)
```

# fc_2_o_agr --> eff step 1
```{r}
# intercept only
mod1e <- lmerTest::lmer(eff ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(mod1e)
lme4::confint.merMod(mod1e)

# adding predictor fc_2_o_agr and time
mod2e <- lmerTest::lmer(eff ~ fc_2_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod2e)
lme4::confint.merMod(mod2e)

# model fit comparison
anova(mod1e, mod2e)
```

# fc_2_o_agr --> eff step 2
```{r}
# adding group mean centered fc_2_o_agr
mod3e <- lmerTest::lmer(eff ~ fc_2_o_agr + gmc_fc_2_o_agr + time + (1|id), US_mlm, REML = FALSE)
summary(mod3e)
lme4::confint.merMod(mod3e)

anova(mod1e, mod2e, mod3e)
```

# fc_2_o_agr --> eff step 3
```{r}
# random slopes time
mod4e <- lmerTest::lmer(eff ~ fc_2_o_agr + gmc_fc_2_o_agr + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(mod4e)
lme4::confint.merMod(mod4e)

# failed to converge
allFit(show.meth.tab = TRUE)
mod4e_all <- allFit(mod4e)

anova(mod1e, mod2e, mod3e, mod4e)
```

# fc_1_o_agr --> behavioural intention
```{r}

```
# fc_2_o_agr --> behavioural intention
```{r}

```

# joy --> behavioural intention
```{r}

```

# eff --> behavioural intention
```{r}

```

# cn, fc & joy --> behavioural intention
```{r}

```

# cn --> joy step 1
```{r}
# intercept only
model1 <- lmerTest::lmer(joy ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(model1)
lme4::confint.merMod(model1)

# adding predictor cn, and covariates secident and time
model2 <- lmerTest::lmer(joy ~ cn + + secident + time + (1|id), US_mlm, REML = FALSE)
summary(model2)
lme4::confint.merMod(model2)

# model fit comparison
anova(model1, model2)
```

# cn --> joy step 2
```{r}
# adding gmc cn and secident
model3 <- lmerTest::lmer(joy ~ cn + gmc_cn + gmc_secident + time + (1|id), US_mlm, REML = FALSE)
summary(model3)
lme4::confint.merMod(model3)

# model fit comparison
anova(model1, model2, model3)
```

# cn --> joy step 3
```{r}
# random slopes time
model4 <- lmerTest::lmer(joy ~ cn + gmc_cn + gmc_secident + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(model4)
lme4::confint.merMod(model4)

# failed to converge
allFit(show.meth.tab = TRUE)
model4_all <- allFit(model4)

anova(model1, model2, model3, model4)
```



# cn --> eff step 1
```{r}
# intercept only
model1a <- lmerTest::lmer(eff ~ 1 + (1|id), US_mlm, REML = FALSE)
summary(model1a)
lme4::confint.merMod(model1a)

# adding predictor cn, and covariates secident and time
model2a <- lmerTest::lmer(eff ~ cn + secident + time + (1|id), US_mlm, REML = FALSE)
summary(model2a)
lme4::confint.merMod(model2a)

# model fit comparison
anova(model1a, model2a)
```

# cn --> eff step 2
```{r}
# adding gmc cn
model3a <- lmerTest::lmer(eff ~ cn + gmc_cn + secident + gmc_secident + time + (1|id), US_mlm, REML = FALSE)
summary(model3a)
lme4::confint.merMod(model3a)

# model fit comparison
anova(model1a, model2a, model3a)
```

# cn --> eff step 3
```{r}
# random slopes time
model4a <- lmerTest::lmer(eff ~ cn + gmc_cn + secident + gmc_secident + time + (1|id) + (0+time|id), US_mlm, REML = FALSE, control = lmerControl(check.nobs.vs.nRE = "ignore"))
summary(model4a)
lme4::confint.merMod(model4a)

# failed to converge
allFit(show.meth.tab = TRUE)
model4a_all <- allFit(model4a)

anova(model1a, model2a, model3a, model4a)
```